---
title: "Covid91 vaccine study Final2024"
author: "Joseph Diaz"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "`r Sys.Date()`" 
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="finalSumr2024.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalDat  %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
```




# Introduction

A new  preventative vaccine (covid921) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

## Males

H₀: Infection rates are the same for vaccine and placebo among males.
H₁: Infection rates differ between vaccine and placebo among males.

```{r}
males <- FinalData %>% filter(sex == "M")

table_males <- table(males$treatment, males$infected)
table_males

# statistical test

fisher.test(table_males)



```
Results for Males

The contingency table for male participants shows:

Placebo group: 262 infected, 9,021 not infected

Vaccine group: 93 infected, 9,016 not infected

Because infection counts are relatively small compared to the total group size (18,392 males), a Fisher’s Exact Test was used.

Since the p-value is far below 0.05, we reject the null hypothesis that infection rates are the same for vaccinated and placebo males.

The odds ratio of 2.82 indicates that placebo males were nearly three times more likely to become infected compared to vaccinated males





## Females

Hypotheses

H₀: Infection rates are the same for vaccine and placebo among females.
H₁: Infection rates differ between vaccine and placebo among females.


```{r}
females <- FinalData %>% filter(sex == "F")

table_females <- table(females$treatment, females$infected)
table_females

fisher.test(table_females)



```

Results for Females

The female subgroup includes 18,328 participants divided between placebo and vaccine. Their infection counts were:

Placebo group: 237 infected, 8,970 not infected

Vaccine group: 128 infected, 8,993 not infected

As with males, infection events are relatively rare, so a Fisher’s Exact Test is the appropriate statistical test.

Since the p-value is far below 0.05, we reject the null hypothesis that infection rates are equal for the vaccinated and placebo female groups.

The odds ratio of 1.86 indicates that placebo females were nearly twice as likely to become infected compared to vaccinated females.



## LGBTQ
Hypotheses

H₀: Infection rates are equal for vaccinated vs. placebo LGBTQ individuals.
H₁: Infection rates are different for LGBTQ individuals.

```{r}

lgbtq <- FinalData %>% filter(LGBTQ == "gay")

table_lgbtq <- table(lgbtq$treatment, lgbtq$infected)
table_lgbtq

fisher.test(table_lgbtq)


```
The  dataset includes 3,060 LGBTQ (“gay”) participants, divided across vaccine and placebo groups. Their infection counts were:

Placebo group: 48 infected, 1,450 not infected

Vaccine group: 12 infected, 1,550 not infected

Because infection events are relatively rare in this subgroup, a Fisher’s Exact Test was used.


The p-value is far below 0.05, so we reject the null hypothesis that infection rates are the same between treatment groups for LGBTQ participants.

The odds ratio of 4.27 means:

LGBTQ individuals in the placebo group were over four times more likely to become infected compared to those receiving the vaccine.





## Druggies


# Overall Results and Conclusions